Case-Base Neural Network: survival analysis with time-varying, higher-order interactions

### 2023-06-30 analyses/plotter.Rmd
updatedArxiv, fixed figure size

library(dplyr)
library(flextable)

@@

@@

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
pdf(file="../StatPaper/figures/Figure2.pdf",width=18,height=10)
g
dev.off()
```

@@

edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
    newgrob <- editGrob(table$grobs[id][[1]], ...)
    table$grobs[id][[1]] <- newgrob
  }
  table
}

@@

@@

  #ft$header$styles$text$bold$data
  ft <- add_header_row(x = ft,top=F, values = c("Method","25%","50%","75%","100%",
                                                "25%","50%","75%","100%"),
                       colwidths = c(1,1,1,1,1,1,1,1,1))
  ft$header$styles$text$bold$data[1,1]<-T
  #ft <- add_header_row(
  #  x = ft, values = c(title),
  #  colwidths = c(9))
  ft <- align(ft, i = c(1,2), part = "header", align = "center")
  ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
  ft<-hline(ft,i=c(1,2),border=bord2,part="header")
  #ft<-hline(ft,i=c(1),border=bord,part="body")
  ft<-fontsize(ft, i = NULL, j = NULL, size =7, part = "all")
  ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
  ft <- align(ft, i = NULL, part = "body", align = "center")
  ft$body$styles$text$font.size$data[,1]<-10

@@

getFlex<-function(result,titler="X: study"){
  Tab<-makeTable(list(result$ipa$data,result$brier$data,result$ibs$data,result$cidx$data,result$auc$data),study="Complex")
  Picks<-getMaxes(Tab)
  Tab<-cleantables(Tab$tab)
  Tab<-convert0(Tab)
  Flex<-flexMaker(tab=Tab,boldPicks = Picks,title=titler)

@@

cFlex<-getFlex(complex,"A: Complex")
mFlex<-getFlex(myeloma,"B: MM")
fFlex<-getFlex(flchain,"C: FLC")
pFlex<-getFlex(prostate,"D: Prostate")

pdf(file="../StatPaper/figures/Table2.pdf",width=14,height=13)
cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
dev.off()

@@

pdf(file="../StatPaper/figures/Table1.pdf",width=12,height=6)
cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
dev.off()

@@

pdf(file="../StatPaper/figures/ComplexPopTime.pdf",width=8,height=6)
a
dev.off()

>>>>

library(dplyr)

@@

library(flextable)

@@

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
pdf(file="../manuscript/figures/Figure2.pdf",width=18,height=10)
g
dev.off()
```

@@

library(flextable)
edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
    newgrob <- editGrob(table$grobs[id][[1]], ...)
    table$grobs[id][[1]] <- newgrob
  }
  table
}

@@

convert1<-function(tab){
  tempNames<-rownames(tab)
  tab<-as.data.frame(lapply(tab, function(x){gsub("1 ","1.00 ", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x){gsub('(1,',"(1.00,", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x) gsub(",1)",",1.00)", x,fixed=T)))
  rownames(tab)<-tempNames
  return(tab)
}

@@

  #ft$header$styles$text$bold$data
  ft <- add_header_row(x = ft,top=F, values = c("Method","25%","50%","75%","100%",
                                                "25%","50%","75%","100%"),
                       colwidths = c(1,1,1,1,1,1,1,1,1))
  ft$header$styles$text$bold$data[1,1]<-T
  #ft <- add_header_row(
  #  x = ft, values = c(title),
  #  colwidths = c(9))
  ft <- flextable::align(ft, i = c(1,2), part = "header", align = "center")
  ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
  ft<-hline(ft,i=c(1,2),border=bord2,part="header")
  #ft<-hline(ft,i=c(1),border=bord,part="body")
  ft<-fontsize(ft, i = NULL, j = NULL, size =12, part = "all")
  ft<-fontsize(ft, i = NULL, j = NULL, size =14, part = "header")
  ft <- flextable::align(ft, i = NULL, part = "body", align = "center")
  ft$body$styles$text$font.size$data[,1]<-14

@@

getFlex<-function(result,titler="X: study"){
  Tab<-makeTable(list(result$ipa$data,result$brier$data,result$ibs$data,result$cidx$data,result$auc$data),study="Complex")
  Picks<-getMaxes(Tab)
  Tab<-cleantables(Tab$tab)
  Tab<-convert0(Tab)
  Tab<-convert1(Tab)
  Flex<-flexMaker(tab=Tab,boldPicks = Picks,title=titler)

@@

cFlex<-getFlex(complex,"A:Complex")
mFlex<-getFlex(myeloma,"B:MM")
fFlex<-getFlex(flchain,"C:FLC")
pFlex<-getFlex(prostate,"D:Prostate")

pdf(file="../manuscript/figures/Table2.pdf",width=19,height=26)
cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
dev.off()

@@

pdf(file="../manuscript/figures/Table1.pdf",width=12,height=6)
cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
dev.off()

@@

pdf(file="../manuscript/figures/ComplexPopTime.pdf",width=8,height=6)
a
dev.off()

### 2023-06-30 manuscript/manuscript.tex
updatedArxiv, fixed figure size

The IPA is a function of time, based on the BS. The IPA score is given by \begin{align}
\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
\end{align} where \(BS_{model}(t)\) represents the BS over time \(t\) for the model of interest and \(BS_{null}(t)\) represents
the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.

\hypertarget{auc}{%
\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$)}\label{auc}}

The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,
\begin{align*}
\resizebox{\textwidth}{!}
{
AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
}
\end{align*}
where \begin{align*}
W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
we can better understand whether a model performs better in terms of calibration or discrimination.

@@

\[
\resizebox{\textwidth}{!}
{
z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
z_{2} \sim \begin{cases}
 N(0,0.5) & \textrm{if } z_{1}=0\\ 
 N(1,0.5) & \textrm{if } z_{1}=1
\end{cases} \qquad \qquad
z_{3} \sim N(1,0.5).
}
\]

@@

\begin{table}[!htbp]
\caption{Performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC)
case study (C) and prostate cancer (Prostate) case study (D). Each table shows performance for each method in each study at $25\%$, $50\%$, $75\%$
and $100\%$ of follow-up time. These tables are included to
provide exact measures of performance. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
DeepHit, DeepSurv, and Optimal (in the complex simulation). The best score at each percent of follow-up time is highlighted in bold. If the average performance
is tied, then all tied values are highlighted.}
\label{tab:megaTable}

\begin{center}\includegraphics[width=1\linewidth]{figures/Table2.pdf} \end{center}

\end{table}

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
 \bibliographystyle{elsarticle-num} 
 \bibliography{cas-refs}

>>>>

The IPA is a function of time, based on the BS. The IPA score is given by \begin{align}
\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
\end{align} where $(BS_{model}(t))$ represents the BS over time $(t)$ for the model of interest and $(BS_{null}(t))$ represents
the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.

\hypertarget{auc2}{%
\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve $AUC_{IPCW}$}\label{auc2}}

The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,

\begin{equation*}
\resizebox{\hsize}{!}{$AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I_{\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})} \cdot I_{T_{i}\leq t,\delta_{i}=1} \cdot \left( 1-I_{T_{j}\leq t,\delta_{j}=1} \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I_{T_{i}\leq t,\delta_{i}=1} \cdot \left( 1-I_{T_{j}\leq t,\delta_{j}=1} \right) \cdot W_{i}(t) \cdot W_{j}(t) },$}
\end{equation*}

where \begin{align*}
W_{i}(t)=\frac{I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} +\frac{I_{T_{i}>t}}{\widehat{G}(t)}.
\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
we can better understand whether a model performs better in terms of calibration or discrimination.

@@

\begin{equation*}
\resizebox{\hsize}{!}{$
z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
z_{2} \sim \begin{cases}
 N(0,0.5) & \textrm{if } z_{1}=0\\ 
 N(1,0.5) & \textrm{if } z_{1}=1
\end{cases} \qquad \qquad
z_{3} \sim N(1,0.5).$}
\end{equation*}

@@

\begin{table}[!htbp]
\caption{Performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC)
case study (C) and prostate cancer (Prostate) case study (D). Each table shows performance for each method in each study at $25\%$, $50\%$, $75\%$
and $100\%$ of follow-up time. These tables are included to
provide exact measures of performance. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
DeepHit, DeepSurv, and Optimal (in the complex simulation). The best score at each percent of follow-up time is highlighted in bold. If the average performance
is tied, then all tied values are highlighted.}
\label{tab:megaTable}
\begin{center}\includegraphics[scale=0.24]{figures/Table2.pdf} \end{center}
\end{table}

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
 \bibliographystyle{elsarticle-num} 
 \bibliography{manuscript}
